\documentclass[a4paper,12pt]{article}
\usepackage{cmap}
\usepackage[cp1251]{inputenc}
\usepackage[english, ukrainian, russian]{babel}
\usepackage[left=1cm,right=1cm,top=1cm,bottom=1cm]{geometry}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{color}

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

\lstset{frame=tb,
  language=Java,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=none,
  numberstyle=\tiny\color{gray},
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=true,
  breakatwhitespace=true,
  tabsize=3
}


\begin{document}

\pagenumbering{gobble}

\begin{large}


\begin{center}
\section*{Сучасні клієнтські веб-технології.}
\end{center}

\medskip
\hrule
\medskip
\medskip

\section*{React.}


\begin{lstlisting}
import './App.css';
import React from 'react';

function App() {
  return (
    <div className='container'>
      <h2>Student list </h2>
      <ul className='stud-list'>
        <li>Ivanov</li>
        <li>Petrov</li>
        <li>Frolov</li>
      </ul>
    </div>
  );
}
\end{lstlisting}



\section*{Лабораторна 1.}

\begin{enumerate}
\item Виберіть один із проектів зі списку нижче.

\item Заімплементуйте бекенд частину, використовуючи фреймворк/технологію по бажанню.

\item Заімплементуйте фронтенд частину аплікації, використовуючи\\ фреймворк/технологію по бажанню.

\item Фронтенд має відсилати http/https запити на бекенд. Заімплементуйте CRUD операції під ці запити.

\item Створіть репозиторій в \url{https://github.com/WebTechCNU} для свого проекту, та загрузіть його туди.

\item Список проектів:

\begin{itemize}
\item Study Buddy: An app that allows students to connect with each other for studying and sharing resources. It could include features like study groups, file sharing, and chat functionality.

\item Task Tracker: A simple app for managing and tracking tasks and assignments. Students can create to-do lists, set reminders, and mark tasks as completed.

\item Time Management Tool: An app that assists in managing productivity time effectively. It could include features like setting work schedules, creating reminders, and tracking time spent on different tasks.

\item Budget Planner: An app that helps users track their expenses, set budgets, and manage their finances. It could include features like expense categorization, monthly summaries, and notifications for overspending.

\item College Marketplace: A platform for students to buy and sell items within their college community. It could include categories for textbooks, furniture, electronics, and other student necessities.

\item College Event Calendar: A centralized platform for students to discover and keep track of events happening on campus, such as workshops, seminars, and social gatherings. It could include event details, RSVP functionality, and reminders.

\item Grade Calculator: An app that helps students calculate their grades based on different assignments, exams, and weightings. It could also provide insights into how different scores would impact their overall grades.

\item Virtual Lab Assistant: An interactive app that simulates laboratory experiments for science and engineering students. It allows them to perform virtual experiments, collect data, and analyze results.

\item Flashcard Generator: An app that allows students to create digital flashcards for studying different subjects. It could include features like importing images, organizing flashcards into decks, and tracking progress.

\item Internship/Job Board: A platform that aggregates internship and job opportunities specifically targeted at students. It provides a centralized hub for students to find and apply for relevant positions based on their interests and qualifications.

\item Book Exchange Platform: A web app that enables users to trade or lend books with other students in their neighborhood. It could include features like book listings, search filters, and messaging capabilities.

\item Student Feedback System: An app that enables students to provide feedback on courses, professors, and other aspects of their educational experience. It could include anonymous feedback options, rating systems, and comment sections.

\item Virtual Career Fair: An app that brings together employers and students in a virtual setting. It allows companies to showcase job/internship opportunities, conduct interviews, and provide resources for career development.

\item Mental Health Support: An app focused on promoting mental health and well-being. It could include features like guided meditation, mood tracking, stress management techniques, and access to mental health resources.

\item Group Project Manager: An app designed to facilitate group projects by helping students assign tasks, set deadlines, and collaborate effectively. It could include features like file sharing, task tracking, and communication tools.

\item Personalized Book Recommendation: Description: Build a book recommendation app that suggests personalized book lists based on the user's reading preferences, ratings, and genres of interest.

\item Food Waste Reduction: Description: Develop an app to reduce food waste by connecting users with nearby restaurants, cafes, or grocery stores offering surplus food at discounted prices before it goes to waste.

\item Dream Journal with AI Analysis: Description: Create a dream journal app where users can record their dreams. The app uses natural language processing and AI to analyze dream patterns and meanings.

\item Music Collaboration Platform: Description: Develop a platform for musicians and artists to collaborate on music projects virtually, allowing them to record, mix, and produce music together.

\item Local Community Exchange Platform: Description: Create a platform where members of a local community can offer and exchange goods, services, and skills with each other, fostering a sense of community and collaboration.

\item Language Learning Game with AI: Description: Build an interactive language learning game that uses AI to adapt to each user's language proficiency and provide personalized lessons and challenges.

\item Real-Time Collaborative Code Editor: Description: Develop a collaborative code editor that allows multiple users to work on the same codebase simultaneously, promoting real-time collaboration among developers.


\end{itemize}
\end{enumerate}



\section*{Лабораторна 2.}

\begin{enumerate}
\item Додайте декілька ендопінтів до вашої бекенд аплікації. Передбачте виклик цих нових ендпоінтів фронтенд аплікацією.

\item Хоча б один із ендпоінтів вашої аплікації має бути GET, і хоча б один POST.

\item Передбачте можливість отримати дані за id в api url одного із ваших ендпоінтів.

Наприклад, $api/people/:id$ ендпоінт, який можна викликати $api/people/1$ та отримати запис із id 1.

\item Додайте базу даних (SQL або no-SQL) до вашої аплікації. В базі мають бути хоча б декілька таблиць/колекцій. Додайте можливість отримання запису одного типу за id іншого. У випадку SQL бази додайте one-to-many або many-to-many зв'язок між хоча б двома таблицями.

Наприклад, якщо ваш проект - це розклад для студентів, додайте типи students та courses, та можливість отримати список курсів, на які записаний студент, або список студентів, записаних на курс.

\item Додайте бекап схеми вашої бази в репозиторій та діаграму бази в документацію репозиторію.

\item Додайте readme.md файл до вашого проекту з чіткими поясненнями, як запустити вашу аплікацію локально.

\item Оформіть завдання данної лабораторної як PR, та надайте посилання на нього в moodle.
\end{enumerate}


\section*{Лабораторна 3.}

\begin{enumerate}
\item Додайте сторінки реєстрації та логіну до вашої фронтенд аплікації.

\item Додайте ендпоінт для реєстрації нового користувача, та ендпоінт для аутентифікації користувача, що вертатиме JWT беарер токен.

\item Додайте перевірку на наявність беарер токена в запиті до хоча б одного із старих ендпоінтів. Передбачте повідомлення помилки 403 для користувача.

\item* \textit{(для бажаючих)} Додайте OAuth2 аутентифікацію до вашої аплікації.

\item Додайте вебсокет до вашої аплікації. Достатньо простого повідомлення з певним інтервалом. При бажанні, можете додати нотифікацію для користувача при певній події (наприклад, інший користувач редагує ресурс, з яким працює даний користувач).

\item* \textit{(для бажаючих)} Додайте graphql квері ендпоінт до бекенд частини аплікації.

\item Оформіть завдання данної лабораторної як PR, та надайте посилання на нього в moodle.
\end{enumerate}




\section*{Лабораторна 4.}

\begin{enumerate}
\item Додайте юніт тести до фронтенд частини вашої аплікації.

\item Проредагуйте actions.yaml файл та додайте перевірку виконання юніт тестів. Додатково: можете додати свої власні кастомні степи в yaml файл.

\item* \textit{(для бажаючих)} Додайте Dockerfile до вашого проекту, та контейнеризуйте ваш додаток.

\item Перевірте ваш додаток на відповідність OWASP10. Додайте фікс на одну із вразливостей із стандарту OWASP (наприклад, низький рівень логування, інєкції, інш.). Опишіть, в чому полягала вразливість, яку ви виправляєте.

\item Оформіть завдання данної лабораторної як PR, та надайте посилання на нього в moodle.
\end{enumerate}


\end{large}

\end{document}

