\documentclass[a4paper,12pt]{article}
\usepackage{cmap}
\usepackage[cp1251]{inputenc}
\usepackage[english, ukrainian, russian]{babel}
\usepackage[left=2cm,right=1.5cm,top=2cm,bottom=2cm]{geometry}
\usepackage{amssymb}

\begin{document}

\pagenumbering{gobble}

\begin{large}

\section*{Combinatorics.}

\begin{enumerate}
\item There are 9 different colored beads. How many necklaces can you select from them?

\item How many games can 19 chess players play among themselves?

\item In how many ways can you choose 5 colors from 17 available?

\item 20 points are marked on the plane (no three points lie on the same line).
How many triangles can be formed with vertices at these points?

\item How many four-digit numbers can be formed from the digits 0, 1, 2, 3, 4, 5, 6, 7
so that the digits of this number are not repeated?

\item How many sequences of "heads" and "tails" can be formed by throwing a coin 10 times?

\item In a football team with 11 men, it is necessary to choose a captain and his goalkeeper. How many ways can it be done?

\item How many different words can be formed by rearranging the letters in the word "infinity"?

\item In how many ways can 25 men be divided into 5 teams of 5 men each?

\item How many permutations of the digits 0, 1, ..., 9 are there, in which the digit 2 occupies the second place, and the number 5 is either the fifth or sixth?


\item a) A beetle crawls along the grid lines from point (0, 0) to point (n, n), at every step increasing each of the coordinates by 1. In how many ways can it achieve its goal?

b) Now the beetle crawls along the lines of the three-dimensional grid from point (0, 0, 0) to the point (n, n, n), increasing each of the coordinates by 1. In how many ways can it achieve its goal?

\item Build the truth table for the following boolean functions. Convert them to DNF and PDNF. Build the logical schema for them. \\
a) $\neg x \vee \neg y \vee (x\wedge y) $; b) $(x\rightarrow y)\rightarrow (y\rightarrow x)$; c) $x\rightarrow (y\rightarrow x)$; \\
d) $(x\rightarrow \neg(y\wedge z)) \rightarrow(z \rightarrow \neg x) $; e) $\neg y \wedge x \wedge (x\rightarrow y)$;
f) $(x\vee \neg x) \rightarrow (y\wedge \neg y)$; \\
g) $(x\rightarrow (x\rightarrow y))\rightarrow (y\sim x) $; h) $(x\rightarrow y)\wedge (y\rightarrow z) $; \\
i) $(x \rightarrow(y \rightarrow z))\wedge (x\rightarrow y)\wedge (x\wedge \neg z) $; \\
j) $(x\wedge y)\rightarrow z $; k) $((x\wedge y)\rightarrow z)\rightarrow((x\rightarrow z) \wedge(y\rightarrow z)) $; \\
l) $ (y\rightarrow (x\rightarrow (y\rightarrow z)))\rightarrow (x\rightarrow z)$;

\end{enumerate}



\end{large}
\end{document}

