\documentclass[11pt,a4paper,sans]{moderncv}
\usepackage[utf8]{inputenc}

\moderncvstyle{classic}                        % style options are 'casual' (default), 'classic', 'oldstyle' and 'banking'
                       % color options 'blue' (default), 'orange', 'green', 'red', 'purple', 'grey' and 'black'


\usepackage[scale=0.87]{geometry}
\setlength{\hintscolumnwidth}{2.5cm}           % if you want to change the width of the column with the dates


\firstname{Vasyl}
\familyname{Melnyk}

\address{Pivdenno-Kyltseva str., 7, apt. 30}{Chernivtsi, Ukraine}
\phone{+380667845191}
\email{windchange7@gmail.com}
\extrainfo{website: }
\renewcommand\refname{Selected publications}
\nopagenumbers{} % uncomment to suppress automatic page numbering for CVs longer than one page


\begin{document}
\maketitle

\section{Education}
\cventry{2015-present}{Ph.D. in Mathematical Analysis}{\newline Chernivtsi National University, Ukraine}{\newline \textit{Subject}: Uniform distances to different functional classes and development of Hahn's theorem on intermediate function}{\newline \textit{Supervisors}: Prof. Volodymyr Maslyuchenko}{}

\cventry{2000-2006}{M.S. in Mathematics}{\newline Chernivtsi National University, Ukraine}{\newline \textit{Subject}: }{\newline \textit{Supervisors}: Prof. Prof. Volodymyr Maslyuchenko}{}

\cventry{2000-2006}{M.S. in Physics}{Neverhood Institute Of Technology}{\newline \textit{Subject}: }{\newline \textit{Supervisors}: Prof. }{}



\section{Research Experience}
\cventry{April 2009 --\\ February 2010 }{Senior Engineer}{Bullshit Institute of Technology}{\newline Study of Horseshit}{}{}



\section{Fellowships and Awards}
\cvline{2009}{Moscow Conference Award (Diploma of excellence)}



\section{Professional Participation}
\cvline{Membership}{Society of Bullshit.}


\section{Computer skills}
\cvitem{OS}{UNIX, Debian GNU/Linux, Windows NT.}




\section{Fields of Scientific Interest}
\cvitem{Image\,processing}{images deconvolution, noise analysis in CCD/CMOS-registered images.}


% Publications from a BibTeX file
\nocite{*}
\bibliographystyle{unsrt}

\bibliography{poopkinbiblio}       % 'publications' is the name of a BibTeX file
\normalsize

\section{Personal information}
% \cventry{Date of birth}{June 20, 1982}{}{Russian Federation}{}{}
\cventry{Languages}{English (upper-intermediate)}{Russian (native), German (basic)}{}{}{}


\end{document} 
