
\begin{center}
ABSTRACT
\end{center}

\textbf{\emph{Melnyk V.} Uniform distances to different functional classes and development of Hahn's theorem on intermediate function.} Manuscript.

\vspace{0.1 cm}

A dissertation for the degree  of Candidate of physical and mathematical sciences (Doctor of philosophy) of speciality 01.01.01 -- "Mathematical analysis" (01 - Mathematics). -- Yuriy Fedkovych Chernivtsi National University, Chernivtsi, 2019.

\vspace{\baselineskip}
The Dissertation is dedicated to investigation of possibility of transfer of the results of Hahn-Dieudonne-Tong-Katetov theorem on intermediate continuous function for two semicontinuous lower and upper respectively functions $g,h:X\rightarrow \mathbb{R}$ such, that $g(x)\leq h(x)$ on $X$ for other classes of functions. Also, uniform distances from functions to different classes of functions.

The Hahn's pairs are introduced in this dissertation, that being the pairs $(g,h)$ of semicontinuous respectively upper and lower functions $g,h:X\rightarrow \mathbb{R}$ on topological space $X$, such, that $g(x)\leq h(x)$ on $X$, and strict Hahn's pairs, when $g(x)<h(x)$ on $X$.

We constructed intermediate affine function for convex and concave functions $g,h:X\rightarrow \mathbb{R}$, $g(x)\leq h(x)$, given on vector space $X$.

Also we have constructed strictly intermediate differentiable functions for strict Hahn's pairs on segments, parallelepipeds in $\mathbb{R}^{n}$, asplund and Hilbert spaces.

With the help of cross topology there were received a number of results on transfer of Hahn's result to the case of separately continuous functions on products of topological spaces, and also for multivalued Hahn's pairs.

Last part of dissertation is dedicated to the problem of finding of uniform distance $d(f,C(X))$ and $d(f,K_{x_{0}}(X))$ from functions, given on topological space $X$, to spaces $C(X)$ of continuous and $K_{x_{0}}(X)$ of quasicontinuous in $x_{0}$ functions.


 \textbf{Key words:}

Hahn's pair, strict Hahn's pair, intermediate function, strictly intermediate function, continuous function, separately continuous maps, asplund spaces, quasicontinuous functions, best approximation, normal spaces, paracompact spaces, partition of unity.

